Real-Time Application of Activity Classification from Body-Worn Accelerometers

### 2016-06-08 notes/paper/paper.tex
Added more to the paper.

The matched filter is based on the cross correlation of the data $\vec{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\vec{x} = \{x\}_{k=0}^{M-1}$.
The cross correlation of $\vec{x}$ with $\vec{y}$ is defined as:
%
\begin{equation} \label{cross_correlation_eq}
(\vec{x} \star \vec{y}) = \left \{\sum_{i=0}^{M}(x_{i} y_{i+k}) \right \}_{k=0}^{N-M-1}
\end{equation}
%
which is simply a moving dot product.
We can then normalize \eqref{cross_correlation_eq} by the following:
%
\begin{equation} \label{norm_cross_correlation_eq}
\widehat{(\vec{x} \star \vec{y})}_k = \frac{(\vec{x} \star \vec{y})_k}{||\vec{x}|| \ || \{y_p\}_{p=k}^{M+k} || }, \quad k = 0,1,...,N-M-1
\end{equation}
%
where $|| \cdot ||$ is the usual $L_2$ norm. \eqref{norm_cross_correlation_eq} is now bound by $[-1,1]$ and has any bias toward high energy removed.
We improve our computational efficiency by applying the convolution theorem to \eqref{cross_correlation_eq} and replacing it with:
%
\begin{equation} \label{conv_theorem}
(\vec{x} \star \vec{y}) = FFT^{-1}(FFT(\vec{x}') FFT(\vec{y})^*)
\end{equation}
%
Where $\vec{x}'$ is $\vec{x}$ that has been zero padded to the length of $\vec{y}$ with the added restriction that the length of $\vec{x}$ is no more than the length of $\vec{y}$ to prevent corruption from circular indexing.
We are most interested where our reference best matches our data.
The resulting max correlation yields our matched-filter output:
%
\begin{equation} \label{matched_filter_eq}
MF(\vec{x},\vec{y}) := \underset{k \in [0, M-N]}{max} \left \{\widehat{(\vec{x} \star \vec{y})}_k \right \}
\end{equation}
%
\section{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time, a user\textquotesingle s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

@@

If $\vec{t} = \{t_k\}_{k=0}^{N_t}$ is our training set over $N_t$ data points, we define a collection of subsets $X$ of $\vec{t}$:
%
\begin{equation} \label{X_subsets_of_training_eq}
X := \left \{ \{t_k\}_{k=p}^{p+M-1} | p=mI \right \}
\end{equation}
%
and $Y$ of $\vec{t}$:
%
\begin{equation} \label{Y_subsets_of_training_eq}
Y := \left \{ \{t_k\}_{k=p}^{p+N-1} | p=nI \right \}
\end{equation}
%
where $M \leq N$, $I$ is a positive, fixed integer, and $(n,m) \in \mathbb{Z}^2$ where $(m,n)$ vary such that $0 \leq (m,n)I$ and $(m,n)I + (M,N) - 1 \leq N_t$. Then we can determine the representative signal $\vec{s} \in X$ for our activity that satisfies the condition:
%
\begin{equation} \label{s_condition}
\sum_{y \in Y}MF(\vec{s}, \vec{y}) \geq \sum_{y \in Y}MF(\vec{x},\vec{y}) \quad \forall \vec{x} \in X
\end{equation}

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
\section{System Hardware}
The sensor used for this analysis is a three-axis accelerometer unit from Kinetisense.
The unit’s acceleration range is -5g to +5g with a sampling rate of 128Hz.
Samples of walking, sitting, running, and bicycling have been recorded with the device and used for the analysis in this paper.
%
\section{Algorithm Software Structure}
%
The algorithms are broken into three main modules: The training, pre-processing of the data used as references for the matched filtering, and the activity classification itself.

>>>>

The matched filter is based on the cross correlation of the data $\vec{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\vec{x} = \{x\}_{k=0}^{M-1}$.
The cross correlation of $\vec{x}$ with $\vec{y}$ is defined as:
%
\begin{equation} \label{cross_correlation_eq}
(\vec{x} \star \vec{y}) = \left \{\sum_{i=0}^{M}x_{i} y_{i+k} \right \}_{k=0}^{N-M-1}
\end{equation}
%
which is simply a moving dot product.
We can then normalize \eqref{cross_correlation_eq} by the following:
%
\begin{equation} \label{norm_cross_correlation_eq}
\widehat{(\vec{x} \star \vec{y})}_k = \frac{(\vec{x} \star \vec{y})_k}{||\vec{x}|| \ || \{y_p\}_{p=k}^{M+k} || }, \quad k = 0,1,...,N-M-1
\end{equation}
%
where $|| \cdot ||$ is the usual $L_2$ norm. \eqref{norm_cross_correlation_eq} is now bound by $[-1,1]$ and has any bias toward high energy removed.
We improve our computational efficiency by applying the convolution theorem to \eqref{cross_correlation_eq} and replacing it with:
%
\begin{equation} \label{conv_theorem}
(\vec{x} \star \vec{y}) = FFT^{-1}(FFT(\vec{x}') FFT(\vec{y})^*)
\end{equation}
%
Where $\vec{x}'$ is $\vec{x}$ that has been zero padded to the length of $\vec{y}$ with the added restriction that the length of $\vec{x}$ is no more than the length of $\vec{y}$.
We are most interested where our reference $\vec{x}$ best matches our data $\vec{y}$.
The resulting max correlation yields our matched-filter for our reference define as:
%
\begin{equation} \label{matched_filter_eq}
MF_{\vec{x}}(\vec{y}) := \underset{k \in [0, M-N]}{max} \left \{\widehat{(\vec{x} \star \vec{y})}_k \right \}
\end{equation}
%
\section{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time, a user\textquotesingle s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

@@

If $\vec{t} = \{t_k\}_{k=0}^{N_t}$ is our training set over $N_t$ data points, we define a collection of subsets $X$ of $\vec{t}$:
%
\begin{equation} \label{X_subsets_of_training_eq}
X := \left \{ \{t_k\}_{k=p}^{p+M-1} | p=mI \right \}
\end{equation}
%
and $Y$ of $\vec{t}$:
%
\begin{equation} \label{Y_subsets_of_training_eq}
Y := \left \{ \{t_k\}_{k=p}^{p+N-1} | p=nI \right \}
\end{equation}
%
where $M \leq N$, $I$ is a positive, fixed integer, and $(n,m) \in \mathbb{Z}^2$ where $(m,n)$ vary such that $0 \leq (m,n)I$ and $(m,n)I + (M,N) - 1 \leq N_t$. Then we can determine the representative signal $\vec{s} \in X$ for our activity that satisfies the condition:
%
\begin{equation} \label{s_condition}
\sum_{y \in Y}MF_{\vec{s}}(\vec{y}) \geq \sum_{y \in Y}MF_{\vec{x}}(\vec{y}) \quad \forall \vec{x} \in X
\end{equation}

We chose this method for its simplicity and robustness against outliers in our training set. We anticipate that the user will have idle motion at the start and end of their training duration when they activate the training mode and with possible disruptions during the process.
%
% Decimation with a tent filter for throughput performance
\section{System Hardware}
The sensor used for this analysis is a three-axis accelerometer unit from Kinetisense.
The unit’s acceleration range is -5g to +5g with a sampling rate of 128Hz.
Samples of walking, sitting, running, and bicycling have been recorded with the device and used for the analysis in this paper.
%
\section{Algorithm Software Structure}
%
The algorithms are broken into three main modules: The training, pre-processing of the data used as references for the matched filtering, and the activity classification itself.
